Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-09-12 paper.tex
Edits Whatever...

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\runtitle{Dynamic Hierarchical Modeling of Sparse Expert Beliefs}

@@

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}},
\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
\and
\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}}

@@

 \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
 \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
 \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}

@@

\begin{abstract}
Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.  This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context in which experts are allowed to update their beliefs at random intervals. The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures. In response to the lack of appropriate methodology, this paper presents a hierarchical model that takes into account the expert's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample. The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 international political events. 
\end{abstract}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. Their probability assessments are often evaluated and compared in regards to \textit{calibration} that measures how closely the frequency of occurrence of the events agree with the given probabilities. For instance, the proportion of occurred events is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts. In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.

Calibration by itself, however, is not sufficient for a useful probability assessment. To see this, consider a relatively stationary process, such as rain on different days in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%. In this setting an expert can assess a constant probability of 0.45 and be well-calibrated. This assessment, however, can be made without any subject-matter expertise or special training. For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information. Therefore a consulting expert should aim to make probability assessments that are as far from the baseline as possible. The extent to which his probabilities differ from the baseline is measured with an important attribute called \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}). If the expert is both sharp and well-calibrated, he is able to forecast the behavior of the process with high certainty and accuracy. From the decision-maker's perspective his advice is highly useful as it involves much information and very little risk. Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}). This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.

One such observation is related to the aggregation of multiple probabilities. There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  results in better predictive performance. Being motivated by the long list of applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many different approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider an investor aiming to assess whether a stock index will finish trading above some threshold on a given date. In order to maximize his overall predictive accuracy, he may consult a group of experts repeatedly over a period of time and adjust his estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice. Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point. As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable to our setting. In response to the lack of appropriate modeling procedures, this paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is highly interpretable and can be used for:
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a subset of 166 geopolitical binary events. After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. An extension of this model to polychotomous outcomes is briefly discussed before model evaluation. The first evaluation section uses synthetic data to study the accuracy to which the two-step procedure is able to estimate parameter values. The second evaluation section illustrates different aspects of the model by applying it to our real-world probability forecasting data. The paper concludes with a discussion on future research directions and model limitations.

@@

The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts. Since a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.

The experts made updates on a very infrequent basis. An expert gave, on average, only 0.016 forecasts per day; that is, a forecast in about every 63 days. The average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the data is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively associated with the self-reported expertise. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, disagree rather strongly, with the disagreement persisting even near the closing dates of the questions.

@@

The hidden state $X_{t,k}$ represents the logit-probability for the $k$th event given all the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. It is reasonable to assume that this probability forecast 
%that makes uses all available information and hence is equivalent to $\logit^{-1}(X_{t,k})$ 
%an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, 
maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
\begin{mydef}
A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
\end{mydef}
\begin{mydef}
A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}). 
\end{mydef}
\noindent
Even though it is unlikely that any single expert has access to all the available information, a large and diverse group of experts may share a considerable portion of the available information. The collective wisdom of the group therefore provides an attractive proxy for $\mathcal{F}_{t,k}$.

Since the experts may believe in false information, hide their true beliefs for ethical bases, or be biased for many other reasons, their probability assessments should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$. In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from $X_{t,k}$. Therefore an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.  Due to the high sparsity of our data, estimating a bias term separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from $X_{t,k}$ is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

 The \textit{random fluctuations} in the hidden process are measured in terms of $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Since each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.

@@

The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model. This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

@@

The goal in this section is to evaluate the ability of the SAC-procedure to capture true parameter values. The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states. The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.

To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 25 experts allocated evenly among five expertise groups. Each expert gives one probability forecast per day with the exception of time $t = 101$, when the event resolves. These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})). In this fashion, the simulation iterates over a three-dimensional grid of parameter values:
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Every grid point is used  40 times to generate a synthetic dataset. The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in. The accuracy of the estimated hidden process and bias vector are measured in  with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively. The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.

@@

\subsection{Incoherent and Imbalanced Data}
\label{practicalMatters}
The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event. For instance, all the 166 questions in our dataset contained both a zero and a one. Transforming such forecasts into the logit-space yields infinities that are highly influential and cause problems in model estimation. To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively. This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).  \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest. Since this implies ignoring a portion of the data, we adopt the first approach by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.  Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.

>>>>

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\runtitle{Dynamic Hierarchical Modeling of Sparse Expert Beliefs}

@@

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
\and
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}

@@

 \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
 \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
 \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}

@@

\begin{abstract}
Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.  This paper advances into unexplored areas of probability aggregation by considering a dynamic context in which experts can update their beliefs at random intervals. The updates occur very infrequently, resulting in a highly sparse dataset that cannot be modeled by standard time-series procedures. In response to the lack of appropriate methodology, this paper presents a hierarchical model that takes into account the expert's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample. The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts over a period of two years on different subsets of 166 international political events. 
\end{abstract}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. Their probability assessments are often evaluated and compared on \textit{calibration}, which measures how closely the frequency of event occurrence agrees with the assigned probabilities. For instance, the proportion of occurrences is 60\% for all those events to which a well-calibrated expert assigned a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts. In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.

Calibration by itself, however, is not sufficient for useful probability estimation. Consider a relatively stationary process, such as rain on different days in a given geographic region, where the observed frequency of occurrence in the last 10 years is 45\%. In this setting an expert could always  assign a constant probability of 0.45 and be well-calibrated. This assessment, however, can be made without any subject-matter expertise or special training. For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information. Therefore experts should aim to make probability assessments that are as far from the baseline as possible. The extent to which their probabilities differ from the baseline is measured by \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}). If the experts are both sharp and well-calibrated, they can forecast the behavior of the process with high certainty and accuracy. From the decision-maker's perspective their advice is highly useful as it involves much information and very little risk. Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}). This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.

%One such observation is related to the aggregation of multiple probabilities. 
There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  improves predictive performance. Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date. To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice. Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point. As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable. This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is highly interpretable and can be used for:
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities, such as expert disagreement and problem difficulty, that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. An extension of this model to polychotomous outcomes is briefly discussed before model evaluation. The first evaluation section uses synthetic data to study how accurately the two-step procedure can estimate parameter values. The second evaluation section applies the model to real-world forecasting data. The paper concludes with a discussion on future research directions and model limitations.

@@

The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period during which the participating experts could update their forecasts as frequently as they liked without penalty. The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts. Since a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.

The experts made updates on a very infrequent basis. An expert gave, on average, only 0.016 forecasts per day; that is, a forecast in about every 63 days. The average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the data are very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively associated with the self-reported expertise. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.

@@

The hidden state $X_{t,k}$ represents the aggregate logit-probability for the $k$th event given all the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. It is reasonable to assume that this probability forecast 
%that makes uses all available information and hence is equivalent to $\logit^{-1}(X_{t,k})$ 
%an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, 
maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
\begin{mydef}
A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
\end{mydef}
\begin{mydef}
A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}). 
\end{mydef}
\noindent
Even though it is unlikely that any single expert has access to all the available information, a large and diverse group of experts may share a considerable portion of the available information. The collective wisdom of the group therefore provides an attractive proxy for $\mathcal{F}_{t,k}$.

Given that the experts may believe in false information, hide their true beliefs, or be biased for many other reasons, their probability assessments should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$. In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from $X_{t,k}$. Therefore an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.  Due to the high sparsity of our data, estimating a bias term separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from $X_{t,k}$ is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

 The \textit{random fluctuations} in the hidden process are measured by $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Since each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.

@@

The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model. This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimizational procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

@@

The goal in this section is to evaluate the ability of the SAC-procedure to capture true parameter values. The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states. The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.

To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 25 experts allocated evenly among five expertise groups. Each expert gives one probability forecast per day with the exception of time $t = 101$, when the event resolves. These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})). In this fashion, the simulation iterates over a three-dimensional grid of parameter values:
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Every grid point is used  40 times to generate a synthetic dataset. The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in. The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively. The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.

@@

\subsection{Incoherent and Imbalanced Data}
\label{practicalMatters}
The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event. For instance, all the 166 questions in our dataset contained both a zero and a one. Transforming such forecasts into the logit-space yields infinities that can cause problems in model estimation. To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively. This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).  \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest. Since this implies ignoring a portion of the data, we adopt an approach similar to \citet{Ariely00theeffects}  by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.  Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.

The second matter is related to the distribution of the class labels in the data. If the set of occurrences is much larger than the set of non-occurrences (or \textit{vice versa}), the dataset is called \textit{imbalanced}. On such data the modeling procedure can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}). Fortunately, it is often possible to use a well-balanced version of the data. The first step is to find a partition $S_0$ and $S_1$ of the question indices $\{1, 2, \dots, K\}$ such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible. This is equivalent to an NP-hard problem known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}). A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g. \citet{kellerer2004knapsack, gent1996phase} for more details on the Partition Problem). After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ for $x = 0$ or $1$. Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator associated with the $k$th question. To define a balanced set of indicators $ \tilde{Z}_k$ for all $k \in S_x$, let
\begin{align*}
 \tilde{Z}_k &= x\\
\tilde{p}_{i,t,k} &=  \begin{cases} 
1-p_{i,t,k}, & \text{if } Z_k = 1-x,\\
p_{i,t,k}, & \text{if } Z_k = x,
\end{cases}
\end{align*}
where $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$. The resulting set $$\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$$ is a balanced version of the data. 
This procedure was used to balance our real-world dataset both in terms of events and time points. The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.
